---
title: 'Demographics factors'
output:
github_document:
toc: yes
toc_float: yes
---

# Helper Functions and T1 demogs

```{r}
helper_func_path = 'https://raw.githubusercontent.com/zenkavi/SRO_Retest_Analyses/master/code/helper_functions/'
eval(parse(text = getURL(paste0(helper_func_path,'get_demographics.R'), ssl.verifypeer = FALSE)))
eval(parse(text = getURL(paste0(helper_func_path,'residualize_baseline.R'), ssl.verifypeer = FALSE)))
eval(parse(text = getURL(paste0(helper_func_path,'find_optimal_components.R'), ssl.verifypeer = FALSE)))
```

## `get_demographics` function

https://github.com/IanEisenberg/Self_Regulation_Ontology/blob/master/selfregulation/utils/utils.py#L241

- Arguments:
  - dataset=None,
  - cleanup=True,
  - num_response_thresh=10,
  - drop_categorical=True, verbose=False

- Procedure:
  - read in `demographic_health.csv`
  - replace `WeightPounds<50` with NA
  - replace `HeightInches<36` with NA
  - replace `CaffienatedSodaCansPerDay<0` with NA
  - replace `CaffieneOtherSourcesDayMG>2000` with NA
  - create `BMI` column `BMI=demogdata['WeightPounds']*0.45 / (demogdata['HeightInches']*0.025)**2`
  - drop categorical vars categorical_vars = ['HispanicLatino','Race',
                        'DiseaseDiagnoses', 'DiseaseDiagnosesOther',
                        'MotivationForParticipation', 'MotivationOther',
                        'NeurologicalDiagnoses',
                        'NeurologicalDiagnosesDescribe',
                        'OtherDebtSources',
                        'OtherDrugs', 'OtherRace', 'OtherTobaccoProducts',
                        'PsychDiagnoses',
                        'PsychDiagnosesOther']
  - create `Obese` column `demogdata['BMI']>30)`
  - only keep variables with fewer NaNs then num_response_thresh

```{r eval=FALSE}
get_demographics = function(dataset=NA, cleanup=TRUE, num_response_thresh=10, drop_categorical=TRUE){
  require(tidyverse)
  
  data = read.csv(dataset)
  
  if(cleanup){
    data = data %>%
      mutate(WeightPounds = ifelse(WeightPounds<50, NA, WeightPounds),
             HeightInches = ifelse(HeightInches<36, NA, HeightInches),
             CaffienatedSodaCansPerDay = ifelse(CaffienatedSodaCansPerDay < 0, NA, CaffienatedSodaCansPerDay),
             CaffieneOtherSourcesDayMG = ifelse(CaffieneOtherSourcesDayMG>2000, NA, CaffieneOtherSourcesDayMG),
             BMI = WeightPounds*0.45/(HeightInches*0.025)^2,
             Obese = ifelse(BMI>30, 1, 0)) %>%
      select(-WeightPounds, -HeightInches)
  }
  
  categorical_vars = c('HispanicLatino','Race',
                        'DiseaseDiagnoses', 'DiseaseDiagnosesOther',
                        'MotivationForParticipation', 'MotivationOther',
                        'NeurologicalDiagnoses',
                        'NeurologicalDiagnosesDescribe',
                        'OtherDebtSources',
                        'OtherDrugs', 'OtherRace', 'OtherTobaccoProducts',
                        'PsychDiagnoses',
                        'PsychDiagnosesOther')
  
  if(drop_categorical){
    data = data %>%
      select(-one_of(categorical_vars))
  }
  
  if(num_response_thresh>0){
    data = data %>%
      select(which(colMeans(is.na(.)) < num_response_thresh/nrow(.)))
  }
  
  return(data)
}
```


```{r}
data_path = '/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/'
release = 'Complete_03-29-2018/'
dataset = 'demographic_health.csv'

demographics_t1 = get_demographics(dataset = paste0(data_path, release, dataset))

demographics_t1
```

## `residualize_baseline` function

https://github.com/IanEisenberg/Self_Regulation_Ontology/blob/master/dimensional_structure/utils.py#L479

- Arguments:
  - df
  - baseline vars: ['Age', 'Sex']

- Procedure:
  - create df with baseline vars extracted from df
  - remove baseline vars from df to create data df
  - impute data df if there is missing data using `fancyimpute` `SimpleFill()`
  - for each variable v in data df regress v ~ Age+Sex
  - new v = v - predicted v using empirical Age+Sex

- Output:
  - data with columns that have the effect of Age and Sex residualized out

```{r eval=FALSE}
residualize_baseline = function(df, baseline_vars = c("Age", "Sex")){
  require(tidyverse)
  
  X = df %>%
    select(one_of(baseline_vars))
  
  Y = df %>%
    select(-one_of(c("X", baseline_vars)))
  
  
  for(v in names(Y)){
    tmp = cbind(X, Y %>% select(v))
    m = lm(as.formula(paste(v,' ~ .')), tmp)
    tmp$v_pred = tmp[,v] - predict(m)
    Y[,v] = tmp$v_pred
  }
  
  return(Y)
  
}
```

```{r}
res_demographics_t1 = residualize_baseline(demographics_t1)
```

## `Demographic Analysis` class

https://github.com/IanEisenberg/Self_Regulation_Ontology/blob/master/dimensional_structure/results.py#L463

- Arguments:
  - subclass of `EFA_Analysis` class
  -  data, residualize=True, residualize_vars=['Age', 'Sex'], boot_iter=1000

- Procedures:
  - calls `residualize_baseline`
  - (for change in retest data also calls `get_demographics` function)
  - `Results` class calls the Demographic Analysis class for its DA attribute with data read is using `get_demographics`
  - `EFA_Analysis` -->
    - `run` -->
      - `get_dimensionality` --> **`find_optimal_components`** --> `psychFA` looping through a range of factor numbers --> returns `best_c` (number of optimal components) and `metrics` df of fit statistic for each number of components
      - `get_loadings` --> if no factor solution has been computed runs `psychFA` with the optimal number of components
      - `computer_higher_order_factors` --> runs `find_optimal_components` and `psychFA` on the correlation matrix of subject level **`psychFA`** scores
      - `get_factor_entropies`

```{r eval=FALSE}
find_optimal_components = function(data, minc=1, maxc=50, fm="ml"){
  
  require(tidyverse)
  
  steps_since_best = 0 # count steps since last best metric.
  maxc = min(maxc, dim(data)[2])
  n_components = seq(minc,maxc)
  metrics = data.frame(comp=n_components, BIC = NA)
  best_metric = NA
  best_c = 0
  
  for(cur_c in n_components){
    out = psych::fa(data, cur_c, rotate='oblimin', fm=fm, scores='tenBerge')
    if(is.null(out$BIC)){
      curr_metric = NA
    }
    else{
      curr_metric = out$BIC
    }
    
    metrics$BIC[which(metrics$comp == cur_c)] = curr_metric
  }
  
  metrics = metrics %>% arrange(BIC)
  
  return(metrics)
}
```

```{r warning=FALSE, message=FALSE}
demog_comp_metrics_t1 = find_optimal_components(res_demographics_t1)

demog_comp_metrics_t1
```

```{r}
demog_fa_t1 = psych::fa(res_demographics_t1, demog_comp_metrics_t1$comp[1], rotate='oblimin', fm='ml', scores='tenBerge')
```

```{r}
demog_fa_loadings_t1 = data.frame(demog_fa_t1$loadings[]) %>%
  mutate(dv = row.names(.)) %>%
  select(dv,everything()) %>%
  rename(Daily_Smoking=ML1,
       Lifetime_Smoking=ML2,
       Problem_Drinking=ML3,
       Mental_Health = ML4,
       Drug_Use = ML5,
       Obesity = ML6,
       Binge_Drinking=ML7,
       Unsafe_Drinking=ML8,
       Income_LifeMilestones=ML9)

demog_fa_loadings_t1
```

```{r}
demog_fa_scores_t1 = data.frame(demog_fa_t1$scores[]) %>%
  mutate(sub_id = demographics_t1$X) %>%
  select(sub_id,everything()) %>%
  rename(Daily_Smoking=ML1,
       Lifetime_Smoking=ML2,
       Problem_Drinking=ML3,
       Mental_Health = ML4,
       Drug_Use = ML5,
       Obesity = ML6,
       Binge_Drinking=ML7,
       Unsafe_Drinking=ML8,
       Income_LifeMilestones=ML9)

demog_fa_scores_t1
```

Compare my factors with Ian's from Python

```{r}
ian_demog = read.csv('/Users/zeynepenkavi/Downloads/demo_factors.csv')
ian_demog = ian_demog %>%
  rename(sub_id = X,
         Drug_Use = Drug.Use,
         Mental_Health = Mental.Health,
         Problem_Drinking = Problem.Drinking,
         Daily_Smoking = Daily.Smoking,
         Binge_Drinking = Binge.Drinking,
         Lifetime_Smoking = Lifetime.Smoking,
         Unsafe_Drinking = Unsafe.Drinking,
         Income_LifeMilestones = Income...Life.Milestones) %>%
  mutate(exec = 'ian')
```

```{r}
demog_fa_scores_t1 %>%
  mutate(exec = "me") %>%
  rbind(.,ian_demog) %>%
  gather(key, value, -sub_id, -exec) %>%
  spread(exec, value) %>%
  ggplot(aes(ian, me))+
  geom_point()+
  geom_abline(intercept=0, slope=1)+
  facet_wrap(~key)
```

# T2 demogs

```{r}
data_path = 'https://raw.githubusercontent.com/zenkavi/Self_Regulation_Ontology/master/Data/'
release = 'Retest_03-29-2018/'
dataset = 'demographic_health.csv'

demographics_t2 = get_demographics(dataset = paste0(data_path, release, dataset))

res_demographics_t2 = residualize_baseline(demographics_t2)
res_demographics_t2 = res_demographics_t2 %>% select(names(res_demographics_t2)[which(names(res_demographics_t2)%in%names(res_demographics_t1))])
```

Predicting factor scores for t2 using t1 model.

```{r error=TRUE}
demog_fa_scores_t2_pred = predict(demog_fa_t1, res_demographics_t2)
```

```{r}
data.frame(demog_fa_scores_t2_pred) %>%
  mutate(sub_id = demographics_t2$X,
         time = "T2") %>%
  select(sub_id,everything()) %>%
  rename(Daily_Smoking=ML1,
       Lifetime_Smoking=ML2,
       Problem_Drinking=ML3,
       Mental_Health = ML4,
       Drug_Use = ML5,
       Obesity = ML6,
       Binge_Drinking=ML7,
       Unsafe_Drinking=ML8,
       Income_LifeMilestones=ML9) %>%
  rbind(demog_fa_scores_t1 %>% 
          filter(sub_id %in% demographics_t2$X) %>%
          mutate(time = "T1")) %>%
  gather(key, value, -sub_id, -time) %>%
  spread(time, value) %>%
  ggplot(aes(T1, T2))+
  geom_point()+
  geom_abline(aes(slope=1, intercept=0))+
  facet_wrap(~key)
```

The optimal number of factors for t2 demographics are smaller than the optimal number for t1 demographics.

```{r warning=FALSE, message=FALSE}
demog_comp_metrics_t2 = find_optimal_components(res_demographics_t2)

demog_comp_metrics_t2
```

Fitting separate factor model for t2 demographics using the optimal number of components.

```{r}
demog_fa_t2 = fa(res_demographics_t2, 4, rotate='oblimin', fm='ml', scores='tenBerge')
```

```{r}
demog_fa_loadings_t2 = data.frame(demog_fa_t2$loadings[])
demog_fa_loadings_t2[abs(demog_fa_loadings_t2)<0.3] = NA
demog_fa_loadings_t2 = demog_fa_loadings_t2 %>%
  mutate(num_load = 4-(is.na(ML1)+is.na(ML2)+is.na(ML3)+is.na(ML4))) %>%
  filter(num_load>0)

demog_fa_loadings_t2 = demog_fa_loadings_t2 %>%
  mutate(dv = row.names(.)) %>%
  select(dv, everything()) %>%
  rename(Smoking = ML1,
         Mental_Health = ML2,
         Drug_Use = ML3,
         Alcohol_Use = ML4)

demog_fa_loadings_t2
```

```{r}
demog_fa_scores_t2 = data.frame(demog_fa_t2$scores[]) %>%
  mutate(sub_id = demographics_t2$X) %>%
  select(sub_id,everything()) %>%
  rename(Smoking = ML1,
         Mental_Health = ML2,
         Drug_Use = ML3,
         Alcohol_Use = ML4)

demog_fa_scores_t2
```