---
title: 'Demographics factors'
output:
github_document:
toc: yes
toc_float: yes
---

## `get_demographics` function

https://github.com/IanEisenberg/Self_Regulation_Ontology/blob/master/selfregulation/utils/utils.py#L241

- Arguments:
  - dataset=None,
  - cleanup=True,
  - num_response_thresh=10,
  - drop_categorical=True, verbose=False

- Procedure:
  - read in `demographic_health.csv`
  - replace `WeightPounds<50` with NA
  - replace `HeightInches<36` with NA
  - replace `CaffienatedSodaCansPerDay<0` with NA
  - replace `CaffieneOtherSourcesDayMG>2000` with NA
  - create `BMI` column `BMI=demogdata['WeightPounds']*0.45 / (demogdata['HeightInches']*0.025)**2`
  - drop categorical vars categorical_vars = ['HispanicLatino','Race',
                        'DiseaseDiagnoses', 'DiseaseDiagnosesOther',
                        'MotivationForParticipation', 'MotivationOther',
                        'NeurologicalDiagnoses',
                        'NeurologicalDiagnosesDescribe',
                        'OtherDebtSources',
                        'OtherDrugs', 'OtherRace', 'OtherTobaccoProducts',
                        'PsychDiagnoses',
                        'PsychDiagnosesOther']
  - create `Obese` column `demogdata['BMI']>30)`
  - only keep variables with fewer NaNs then num_response_thresh

- Output:

```{r}
get_demographics = function(dataset=NA, cleanup=TRUE, num_response_thresh=10, drop_categorical=TRUE){
  require(tidyverse)
  
  data = read.csv(dataset)
  
  if(cleanup){
    data = data %>%
      mutate(WeightPounds = ifelse(WeightPounds<50, NA, WeightPounds),
             HeightInches = ifelse(HeightInches<36, NA, HeightInches),
             CaffienatedSodaCansPerDay = ifelse(CaffienatedSodaCansPerDay < 0, NA, CaffienatedSodaCansPerDay),
             CaffieneOtherSourcesDayMG = ifelse(CaffieneOtherSourcesDayMG>2000, NA, CaffieneOtherSourcesDayMG),
             BMI = WeightPounds*0.45/(HeightInches*0.025)^2,
             Obese = ifelse(BMI>30, 1, 0))
  }
  
  categorical_vars = c('HispanicLatino','Race',
                        'DiseaseDiagnoses', 'DiseaseDiagnosesOther',
                        'MotivationForParticipation', 'MotivationOther',
                        'NeurologicalDiagnoses',
                        'NeurologicalDiagnosesDescribe',
                        'OtherDebtSources',
                        'OtherDrugs', 'OtherRace', 'OtherTobaccoProducts',
                        'PsychDiagnoses',
                        'PsychDiagnosesOther')
  
  if(drop_categorical){
    data = data %>%
      select(-one_of(categorical_vars))
  }
  
  if(num_response_thresh>0){
    data = data %>%
      select(which(colMeans(is.na(.)) < num_response_thresh/nrow(.)))
  }
  
  return(data)
}
```


```{r}
data_path = '/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/'
release = 'Complete_03-29-2018/'
dataset = 'demographic_health.csv'

demographics = get_demographics(dataset = paste0(data_path, release, dataset))

demographics
```

## `residualize_baseline` function

https://github.com/IanEisenberg/Self_Regulation_Ontology/blob/master/dimensional_structure/utils.py#L479

- Arguments:
  - df
  - baseline vars: ['Age', 'Sex']

- Procedure:
  - create df with baseline vars extracted from df
  - remove baseline vars from df to create data df
  - impute data df if there is missing data using `fancyimpute` `SimpleFill()`
  - for each variable v in data df regress v ~ Age+Sex
  - new v = v - predicted v using empirical Age+Sex

- Output:
  - data with columns that have the effect of Age and Sex residualized out

```{r}

```

## `Demographic Analysis` class

https://github.com/IanEisenberg/Self_Regulation_Ontology/blob/master/dimensional_structure/results.py#L463

- Arguments:
  - subclass of `EFA_Analysis` class
  -  data, residualize=True, residualize_vars=['Age', 'Sex'], boot_iter=1000

- Procedures:
  - calls `residualize_baseline`
  - (for change in retest data also calls `get_demographics` function)
  - `Results` class calls the Demographic Analysis class for its DA attribute with data read is using `get_demographics`
  - `EFA_Analysis` -->
    - `run` -->
      - `get_dimensionality` --> **`find_optimal_components`** --> `psychFA` looping through a range of factor numbers --> returns `best_c` (number of optimal components) and `metrics` df of fit statistic for each number of components
      - `get_loadings` --> if no factor solution has been computed runs `psychFA` with the optimal number of components
      - `computer_higher_order_factors` --> runs `find_optimal_components` and `psychFA` on the correlation matrix of subject level **`psychFA`** scores
      - `get_factor_entropies`

```{r eval=FALSE}
psych::fa(data, n_components, rotate='oblimin', fm='ml', n_obs=0,
                      scores='tenBerge')
```

- Output:

```{r}

```
