---
title: 'Reliability of latent variables'
output:
github_document:
toc: yes
toc_float: yes
---

```{r}
ddm_workspace_scripts = '/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_DDM_Analyses/code/workspace_scripts/'
source(paste0(ddm_workspace_scripts,'ez_fa_data.R'))

fig_path = '/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_DDM_Analyses/output/figures/'

library(lavaan)
library(semTools)
require(tidyverse)
theme_set(theme_bw())
options(scipen = 1, digits = 4)
cbbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
theme_set(theme_bw())
ggplot <- function(...) ggplot2::ggplot(...) + scale_fill_manual(values=cbbPalette) + scale_color_manual(values=cbbPalette)

```

In a [previous notebook](ttps://zenkavi.github.io/SRO_DDM_Analyses/output/reports/DimReductionComparison.nb.html) I was stuck on how to evaluate the reliability of a factor solution for data measures at two time points to check whether lower dimensional projections are stable enough across time to be used in trait-level analyses. 

I tried two different approaches to this:  

1. Predicting factor scores for T2 data using the factor solution from T1. I learned that there are different ways of calculating factor scores and only some of these yield reliable factor scores.  

2. Fitting a separate 3 factor model to T2 data and comparing factor scores and factor loadings to each other. Though the solutions were very similar to each other the model being compared is not the same across time points in this approach.

Dave suggested: 
>The standard way to evaluate test-retest correlations with latent variables is to fit a latent variable for each time and then allow the latent variables to correlate. That is the test-retest correlation for the best, “true” score measure at each time point. 

Based on this I did some reading (particularly important were slides 87-91 of [this presentation](http://users.ugent.be/~yrosseel/lavaan/zurich2017/INTRO/lavaan_intro_zurich2017.pdf) and slides 32-41 of [this presentation](http://www.personality-project.org/revelle/syllabi/454/rosseel_sem_longitudinal.pdf)) and here I impelement two ways of modeling the data from the two points together using the time point as a latent variable.  

```{r}
drift_vars = (measure_labels %>% filter(raw_fit == "EZ" & rt_acc == "drift rate") %>% select(dv))[,]
thresh_vars = (measure_labels %>% filter(raw_fit == "EZ" & rt_acc == "threshold") %>% select(dv))[,]
non_dec_vars = (measure_labels %>% filter(raw_fit == "EZ" & rt_acc == "non-decision") %>% select(dv))[,]

drift_vars_sub = (measure_labels %>% filter(raw_fit == "EZ" & rt_acc == "drift rate" & overall_difference != "condition") %>% select(dv))[,]
thresh_vars_sub = (measure_labels %>% filter(raw_fit == "EZ" & rt_acc == "threshold" & overall_difference != "condition") %>% select(dv))[,]
non_dec_vars_sub = (measure_labels %>% filter(raw_fit == "EZ" & rt_acc == "non-decision" & overall_difference != "condition") %>% select(dv))[,]
```

# Time as grouping variable

In this first approach ...

```{r}
tmp1 = res_clean_test_data_ez %>% select(drift_vars, thresh_vars, non_dec_vars)
tmp1$time = "t1"

tmp2 = res_clean_retest_data_ez %>% select(drift_vars, thresh_vars, non_dec_vars)
tmp2$time = "t2"

group_time_data = rbind(tmp1, tmp2)
```

```{r}
#What the paste statement result should look like
ez.group.time.model <-'
drift =~ adaptive_n_back.EZ_drift + attention_network_task.EZ_drift + attention_network_task.EZ_drift_congruent + attention_network_task.EZ_drift_incongruent + attention_network_task.EZ_drift_neutral + attention_network_task.alerting_EZ_drift + attention_network_task.conflict_EZ_drift + attention_network_task.orienting_EZ_drift + choice_reaction_time.EZ_drift + directed_forgetting.EZ_drift + directed_forgetting.EZ_drift_con + directed_forgetting.EZ_drift_neg + directed_forgetting.EZ_drift_pos + directed_forgetting.proactive_interference_EZ_drift + dot_pattern_expectancy.AY.BY_EZ_drift + dot_pattern_expectancy.BX.BY_EZ_drift + dot_pattern_expectancy.EZ_drift + dot_pattern_expectancy.EZ_drift_AX + dot_pattern_expectancy.EZ_drift_AY + dot_pattern_expectancy.EZ_drift_BX + dot_pattern_expectancy.EZ_drift_BY + local_global_letter.EZ_drift + local_global_letter.EZ_drift_congruent + local_global_letter.EZ_drift_global + local_global_letter.EZ_drift_incongruent + local_global_letter.EZ_drift_local + local_global_letter.EZ_drift_neutral + local_global_letter.EZ_drift_stay + local_global_letter.EZ_drift_switch + local_global_letter.conflict_EZ_drift + local_global_letter.global_bias_EZ_drift + local_global_letter.switch_cost_EZ_drift + motor_selective_stop_signal.EZ_drift + recent_probes.EZ_drift + recent_probes.EZ_drift_rec_neg + recent_probes.EZ_drift_rec_pos + recent_probes.EZ_drift_xrec + recent_probes.EZ_drift_xrec_neg + recent_probes.EZ_drift_xrec_pos + recent_probes.proactive_interference_EZ_drift + shape_matching.EZ_drift + shape_matching.EZ_drift_DDD + shape_matching.EZ_drift_DDS + shape_matching.EZ_drift_DNN + shape_matching.EZ_drift_DSD + shape_matching.EZ_drift_SDD + shape_matching.EZ_drift_SNN + shape_matching.EZ_drift_SSS + shape_matching.stimulus_interference_EZ_drift + simon.EZ_drift + simon.EZ_drift_congruent + simon.EZ_drift_incongruent + simon.simon_EZ_drift + stim_selective_stop_signal.EZ_drift + stop_signal.EZ_drift + stroop.EZ_drift + stroop.EZ_drift_congruent + stroop.EZ_drift_incongruent + stroop.stroop_EZ_drift + threebytwo.EZ_drift + threebytwo.EZ_drift_cue_stay_100.0 + threebytwo.EZ_drift_cue_stay_900.0 + threebytwo.EZ_drift_cue_switch_100.0 + threebytwo.EZ_drift_cue_switch_900.0 + threebytwo.EZ_drift_task_switch_100.0 + threebytwo.EZ_drift_task_switch_900.0 + threebytwo.cue_switch_cost_EZ_drift_100.0 + threebytwo.cue_switch_cost_EZ_drift_900.0 + threebytwo.task_switch_cost_EZ_drift_100.0 + threebytwo.task_switch_cost_EZ_drift_900.0
thresh =~ adaptive_n_back.EZ_thresh + attention_network_task.EZ_thresh + attention_network_task.EZ_thresh_congruent + attention_network_task.EZ_thresh_incongruent + attention_network_task.EZ_thresh_neutral + attention_network_task.alerting_EZ_thresh + attention_network_task.conflict_EZ_thresh + attention_network_task.orienting_EZ_thresh + choice_reaction_time.EZ_thresh + directed_forgetting.EZ_thresh + directed_forgetting.EZ_thresh_con + directed_forgetting.EZ_thresh_neg + directed_forgetting.EZ_thresh_pos + directed_forgetting.proactive_interference_EZ_thresh + dot_pattern_expectancy.AY.BY_EZ_thresh + dot_pattern_expectancy.BX.BY_EZ_thresh + dot_pattern_expectancy.EZ_thresh + dot_pattern_expectancy.EZ_thresh_AX + dot_pattern_expectancy.EZ_thresh_AY + dot_pattern_expectancy.EZ_thresh_BX + dot_pattern_expectancy.EZ_thresh_BY + local_global_letter.EZ_thresh + local_global_letter.EZ_thresh_congruent + local_global_letter.EZ_thresh_global + local_global_letter.EZ_thresh_incongruent + local_global_letter.EZ_thresh_local + local_global_letter.EZ_thresh_neutral + local_global_letter.EZ_thresh_stay + local_global_letter.EZ_thresh_switch + local_global_letter.conflict_EZ_thresh + local_global_letter.global_bias_EZ_thresh + local_global_letter.switch_cost_EZ_thresh + motor_selective_stop_signal.EZ_thresh + recent_probes.EZ_thresh + recent_probes.EZ_thresh_rec_neg + recent_probes.EZ_thresh_rec_pos + recent_probes.EZ_thresh_xrec + recent_probes.EZ_thresh_xrec_neg + recent_probes.EZ_thresh_xrec_pos + recent_probes.proactive_interference_EZ_thresh + shape_matching.EZ_thresh + shape_matching.EZ_thresh_DDD + shape_matching.EZ_thresh_DDS + shape_matching.EZ_thresh_DNN + shape_matching.EZ_thresh_DSD + shape_matching.EZ_thresh_SDD + shape_matching.EZ_thresh_SNN + shape_matching.EZ_thresh_SSS + shape_matching.stimulus_interference_EZ_thresh + simon.EZ_thresh + simon.EZ_thresh_congruent + simon.EZ_thresh_incongruent + simon.simon_EZ_thresh + stim_selective_stop_signal.EZ_thresh + stop_signal.EZ_thresh + stroop.EZ_thresh + stroop.EZ_thresh_congruent + stroop.EZ_thresh_incongruent + stroop.stroop_EZ_thresh + threebytwo.EZ_thresh + threebytwo.EZ_thresh_cue_stay_100.0 + threebytwo.EZ_thresh_cue_stay_900.0 + threebytwo.EZ_thresh_cue_switch_100.0 + threebytwo.EZ_thresh_cue_switch_900.0 + threebytwo.EZ_thresh_task_switch_100.0 + threebytwo.EZ_thresh_task_switch_900.0 + threebytwo.cue_switch_cost_EZ_thresh_100.0 + threebytwo.cue_switch_cost_EZ_thresh_900.0 + threebytwo.task_switch_cost_EZ_thresh_100.0 + threebytwo.task_switch_cost_EZ_thresh_900.0
non_dec =~adaptive_n_back.EZ_non_decision + attention_network_task.EZ_non_decision + attention_network_task.EZ_non_decision_congruent + attention_network_task.EZ_non_decision_incongruent + attention_network_task.EZ_non_decision_neutral + attention_network_task.alerting_EZ_non_decision + attention_network_task.conflict_EZ_non_decision + attention_network_task.orienting_EZ_non_decision + choice_reaction_time.EZ_non_decision + directed_forgetting.EZ_non_decision + directed_forgetting.EZ_non_decision_con + directed_forgetting.EZ_non_decision_neg + directed_forgetting.EZ_non_decision_pos + directed_forgetting.proactive_interference_EZ_non_decision + dot_pattern_expectancy.AY.BY_EZ_non_decision + dot_pattern_expectancy.BX.BY_EZ_non_decision + dot_pattern_expectancy.EZ_non_decision + dot_pattern_expectancy.EZ_non_decision_AX + dot_pattern_expectancy.EZ_non_decision_AY + dot_pattern_expectancy.EZ_non_decision_BX + dot_pattern_expectancy.EZ_non_decision_BY + local_global_letter.EZ_non_decision + local_global_letter.EZ_non_decision_congruent + local_global_letter.EZ_non_decision_global + local_global_letter.EZ_non_decision_incongruent + local_global_letter.EZ_non_decision_local + local_global_letter.EZ_non_decision_neutral + local_global_letter.EZ_non_decision_stay + local_global_letter.EZ_non_decision_switch + local_global_letter.conflict_EZ_non_decision + local_global_letter.global_bias_EZ_non_decision + local_global_letter.switch_cost_EZ_non_decision + motor_selective_stop_signal.EZ_non_decision + recent_probes.EZ_non_decision + recent_probes.EZ_non_decision_rec_neg + recent_probes.EZ_non_decision_rec_pos + recent_probes.EZ_non_decision_xrec + recent_probes.EZ_non_decision_xrec_neg + recent_probes.EZ_non_decision_xrec_pos + recent_probes.proactive_interference_EZ_non_decision + shape_matching.EZ_non_decision + shape_matching.EZ_non_decision_DDD + shape_matching.EZ_non_decision_DDS + shape_matching.EZ_non_decision_DNN + shape_matching.EZ_non_decision_DSD + shape_matching.EZ_non_decision_SDD + shape_matching.EZ_non_decision_SNN + shape_matching.EZ_non_decision_SSS + shape_matching.stimulus_interference_EZ_non_decision + simon.EZ_non_decision + simon.EZ_non_decision_congruent + simon.EZ_non_decision_incongruent + simon.simon_EZ_non_decision + stim_selective_stop_signal.EZ_non_decision + stop_signal.EZ_non_decision + stroop.EZ_non_decision + stroop.EZ_non_decision_congruent + stroop.EZ_non_decision_incongruent + stroop.stroop_EZ_non_decision + threebytwo.EZ_non_decision + threebytwo.EZ_non_decision_cue_stay_100.0 + threebytwo.EZ_non_decision_cue_stay_900.0 + threebytwo.EZ_non_decision_cue_switch_100.0 + threebytwo.EZ_non_decision_cue_switch_900.0 + threebytwo.EZ_non_decision_task_switch_100.0 + threebytwo.EZ_non_decision_task_switch_900.0 + threebytwo.cue_switch_cost_EZ_non_decision_100.0 + threebytwo.cue_switch_cost_EZ_non_decision_900.0 + threebytwo.task_switch_cost_EZ_non_decision_100.0 + threebytwo.task_switch_cost_EZ_non_decision_900.0'
```

```{r}
ez.group.time.model <- paste0("\ndrift=~", paste0(drift_vars, collapse= " + "),
       "\nthresh=~", paste0(thresh_vars, collapse= " + "),
       "\nnon_dec=~", paste0(non_dec_vars, collapse= " + "),
        collapse = "\n ")
```

```{r}
tmp1 = res_clean_test_data_ez %>% select(drift_vars_sub, thresh_vars_sub, non_dec_vars_sub)
tmp1$time = "t1"

tmp2 = res_clean_retest_data_ez %>% select(drift_vars_sub, thresh_vars_sub, non_dec_vars_sub)
tmp2$time = "t2"

group_time_data_sub = rbind(tmp1, tmp2)
```

```{r}
#What the paste statement result should look like
ez.group.time.model.sub <-'
drift =~ adaptive_n_back.EZ_drift + attention_network_task.EZ_drift +  attention_network_task.alerting_EZ_drift + attention_network_task.conflict_EZ_drift + attention_network_task.orienting_EZ_drift + choice_reaction_time.EZ_drift + directed_forgetting.EZ_drift + directed_forgetting.proactive_interference_EZ_drift + dot_pattern_expectancy.AY.BY_EZ_drift + dot_pattern_expectancy.BX.BY_EZ_drift + dot_pattern_expectancy.EZ_drift + local_global_letter.EZ_drift + local_global_letter.conflict_EZ_drift + local_global_letter.global_bias_EZ_drift + local_global_letter.switch_cost_EZ_drift + motor_selective_stop_signal.EZ_drift + recent_probes.EZ_drift + recent_probes.proactive_interference_EZ_drift + shape_matching.EZ_drift + shape_matching.stimulus_interference_EZ_drift + simon.EZ_drift + simon.simon_EZ_drift + stim_selective_stop_signal.EZ_drift + stop_signal.EZ_drift + stroop.EZ_drift + stroop.stroop_EZ_drift + threebytwo.EZ_drift + threebytwo.cue_switch_cost_EZ_drift_100.0 + threebytwo.cue_switch_cost_EZ_drift_900.0 + threebytwo.task_switch_cost_EZ_drift_100.0 + threebytwo.task_switch_cost_EZ_drift_900.0

thresh =~ adaptive_n_back.EZ_thresh + attention_network_task.EZ_thresh + attention_network_task.alerting_EZ_thresh + attention_network_task.conflict_EZ_thresh + attention_network_task.orienting_EZ_thresh + choice_reaction_time.EZ_thresh + directed_forgetting.EZ_thresh +  directed_forgetting.proactive_interference_EZ_thresh + dot_pattern_expectancy.AY.BY_EZ_thresh + dot_pattern_expectancy.BX.BY_EZ_thresh + dot_pattern_expectancy.EZ_thresh + local_global_letter.EZ_thresh +  local_global_letter.conflict_EZ_thresh + local_global_letter.global_bias_EZ_thresh + local_global_letter.switch_cost_EZ_thresh + motor_selective_stop_signal.EZ_thresh + recent_probes.EZ_thresh + recent_probes.proactive_interference_EZ_thresh + shape_matching.EZ_thresh + shape_matching.stimulus_interference_EZ_thresh + simon.EZ_thresh + simon.simon_EZ_thresh + stim_selective_stop_signal.EZ_thresh + stop_signal.EZ_thresh + stroop.EZ_thresh + stroop.stroop_EZ_thresh + threebytwo.EZ_thresh + threebytwo.cue_switch_cost_EZ_thresh_100.0 + threebytwo.cue_switch_cost_EZ_thresh_900.0 + threebytwo.task_switch_cost_EZ_thresh_100.0 + threebytwo.task_switch_cost_EZ_thresh_900.0

non_dec =~adaptive_n_back.EZ_non_decision + attention_network_task.EZ_non_decision + attention_network_task.alerting_EZ_non_decision + attention_network_task.conflict_EZ_non_decision + attention_network_task.orienting_EZ_non_decision + choice_reaction_time.EZ_non_decision + directed_forgetting.EZ_non_decision + directed_forgetting.proactive_interference_EZ_non_decision + dot_pattern_expectancy.AY.BY_EZ_non_decision + dot_pattern_expectancy.BX.BY_EZ_non_decision + dot_pattern_expectancy.EZ_non_decision + local_global_letter.EZ_non_decision + local_global_letter.switch_cost_EZ_non_decision + motor_selective_stop_signal.EZ_non_decision + recent_probes.EZ_non_decision + recent_probes.proactive_interference_EZ_non_decision + shape_matching.EZ_non_decision + shape_matching.stimulus_interference_EZ_non_decision + simon.EZ_non_decision + simon.simon_EZ_non_decision + stim_selective_stop_signal.EZ_non_decision + stop_signal.EZ_non_decision + stroop.EZ_non_decision + stroop.stroop_EZ_non_decision + threebytwo.EZ_non_decision + threebytwo.cue_switch_cost_EZ_non_decision_100.0 + threebytwo.cue_switch_cost_EZ_non_decision_900.0 + threebytwo.task_switch_cost_EZ_non_decision_100.0 + threebytwo.task_switch_cost_EZ_non_decision_900.0'
```

```{r}
ez.group.time.model.sub <- paste0("\ndrift=~", paste0(drift_vars_sub, collapse= " + "),
       "\nthresh=~", paste0(thresh_vars_sub, collapse= " + "),
       "\nnon_dec=~", paste0(non_dec_vars_sub, collapse= " + "),
        collapse = "\n ")
```

```{r}
ez.group.time.model.sub.fit = cfa(ez.group.time.model.sub, data = group_time_data_sub, group = "time")
```

```{r}
summary(ez.group.time.model.sub.fit, fit.measures=TRUE)
```

```{r}
measurementInvariance(ez.group.time.model.sub, data = group_time_data, group = "time", strict=FALSE)
```

See slide #33 [here](http://www.personality-project.org/revelle/syllabi/454/rosseel_sem_longitudinal.pdf) for why this approach of using time as a grouping variable is tempting BUT not the correct way of assesing whether the means of latent variables change significantly

```{r}
#Strong invariance model
ez.group.time.model.fit_si = cfa(ez.group.time.model.sub, data = group_time_data, group = "time",  group.equal = c("loadings", "intercepts"))
```

# Time as latent variable

Based on the concern raised above I tried checking whether the means of two separate factors (e.g. one consisting of all the drift rate variables from  T1 and another factor consisting of all the drift rate variables from T2) differed from each other. 

```{r}
# tmp1 = res_clean_test_data_ez %>% select(drift_vars_sub, thresh_vars_sub, non_dec_vars_sub)
tmp1 = res_clean_test_data_ez %>% select(drift_vars_sub, thresh_vars_sub)
names(tmp1) = paste0(names(tmp1), "_1")
# tmp2 = res_clean_retest_data_ez %>% select(drift_vars_sub, thresh_vars_sub, non_dec_vars_sub)
tmp2 = res_clean_retest_data_ez %>% select(drift_vars_sub, thresh_vars_sub)
names(tmp2) = paste0(names(tmp2), "_2")
latent_time_data = cbind(tmp1, tmp2)

MEAN <- c(apply(tmp1, 2, mean),apply(tmp2, 2, mean)) 

SDS <- c(apply(tmp1, 2, sd),apply(tmp2, 2, sd)) 

x = cor(latent_time_data)
x = cor.smooth(x)
lower = x[lower.tri(x, diag = T)]

COV <- getCov(lower, sds=SDS, names = names(latent_time_data))
```

```{r eval=FALSE}
#What the resulting model string should look like
ez.latent.time.model <- '
drift_1 =~ adaptive_n_back.EZ_drift_1 + attention_network_task.EZ_drift_1 + attention_network_task.alerting_EZ_drift_1 + attention_network_task.conflict_EZ_drift_1 + attention_network_task.orienting_EZ_drift_1 + choice_reaction_time.EZ_drift_1 + directed_forgetting.EZ_drift_1 + directed_forgetting.proactive_interference_EZ_drift_1 + dot_pattern_expectancy.AY.BY_EZ_drift_1 + dot_pattern_expectancy.BX.BY_EZ_drift_1 + dot_pattern_expectancy.EZ_drift_1 + local_global_letter.EZ_drift_1 + local_global_letter.conflict_EZ_drift_1 + local_global_letter.global_bias_EZ_drift_1 + local_global_letter.switch_cost_EZ_drift_1 + motor_selective_stop_signal.EZ_drift_1 + recent_probes.EZ_drift_1 + recent_probes.proactive_interference_EZ_drift_1 + shape_matching.EZ_drift_1 + shape_matching.stimulus_interference_EZ_drift_1 + simon.EZ_drift_1 + simon.simon_EZ_drift_1 + stim_selective_stop_signal.EZ_drift_1 + stop_signal.EZ_drift_1 + stroop.EZ_drift_1 + stroop.stroop_EZ_drift_1 + threebytwo.EZ_drift_1 + threebytwo.cue_switch_cost_EZ_drift_100.0_1 + threebytwo.cue_switch_cost_EZ_drift_900.0_1 + threebytwo.task_switch_cost_EZ_drift_100.0_1 + threebytwo.task_switch_cost_EZ_drift_900.0_1
thresh_1 =~ adaptive_n_back.EZ_thresh_1 + attention_network_task.EZ_thresh_1 + attention_network_task.alerting_EZ_thresh_1 + attention_network_task.conflict_EZ_thresh_1 + attention_network_task.orienting_EZ_thresh_1 + choice_reaction_time.EZ_thresh_1 + directed_forgetting.EZ_thresh_1 + directed_forgetting.proactive_interference_EZ_thresh_1 + dot_pattern_expectancy.AY.BY_EZ_thresh_1 + dot_pattern_expectancy.BX.BY_EZ_thresh_1 + dot_pattern_expectancy.EZ_thresh_1 + local_global_letter.EZ_thresh_1 + local_global_letter.conflict_EZ_thresh_1 + local_global_letter.global_bias_EZ_thresh_1 + local_global_letter.switch_cost_EZ_thresh_1 + motor_selective_stop_signal.EZ_thresh_1 + recent_probes.EZ_thresh_1 + recent_probes.proactive_interference_EZ_thresh_1 + shape_matching.EZ_thresh_1 + shape_matching.stimulus_interference_EZ_thresh_1 + simon.EZ_thresh_1 + simon.simon_EZ_thresh_1 + stim_selective_stop_signal.EZ_thresh_1 + stop_signal.EZ_thresh_1 + stroop.EZ_thresh_1 + stroop.stroop_EZ_thresh_1 + threebytwo.EZ_thresh_1 + threebytwo.cue_switch_cost_EZ_thresh_100.0_1 + threebytwo.cue_switch_cost_EZ_thresh_900.0_1 + threebytwo.task_switch_cost_EZ_thresh_100.0_1 + threebytwo.task_switch_cost_EZ_thresh_900.0_1
non_dec_1 =~ adaptive_n_back.EZ_non_decision_1 + attention_network_task.EZ_non_decision_1 + attention_network_task.alerting_EZ_non_decision_1 + attention_network_task.conflict_EZ_non_decision_1 + attention_network_task.orienting_EZ_non_decision_1 + choice_reaction_time.EZ_non_decision_1 + directed_forgetting.EZ_non_decision_1 + directed_forgetting.proactive_interference_EZ_non_decision_1 + dot_pattern_expectancy.AY.BY_EZ_non_decision_1 + dot_pattern_expectancy.BX.BY_EZ_non_decision_1 + dot_pattern_expectancy.EZ_non_decision_1 + local_global_letter.EZ_non_decision_1 + local_global_letter.switch_cost_EZ_non_decision_1 + motor_selective_stop_signal.EZ_non_decision_1 + recent_probes.EZ_non_decision_1 + recent_probes.proactive_interference_EZ_non_decision_1 + shape_matching.EZ_non_decision_1 + shape_matching.stimulus_interference_EZ_non_decision_1 + simon.EZ_non_decision_1 + simon.simon_EZ_non_decision_1 + stim_selective_stop_signal.EZ_non_decision_1 + stop_signal.EZ_non_decision_1 + stroop.EZ_non_decision_1 + stroop.stroop_EZ_non_decision_1 + threebytwo.EZ_non_decision_1 + threebytwo.cue_switch_cost_EZ_non_decision_100.0_1 + threebytwo.cue_switch_cost_EZ_non_decision_900.0_1 + threebytwo.task_switch_cost_EZ_non_decision_100.0_1 + threebytwo.task_switch_cost_EZ_non_decision_900.0_1
drift_2 =~ adaptive_n_back.EZ_drift_2 + attention_network_task.EZ_drift_2 + attention_network_task.alerting_EZ_drift_2 + attention_network_task.conflict_EZ_drift_2 + attention_network_task.orienting_EZ_drift_2 + choice_reaction_time.EZ_drift_2 + directed_forgetting.EZ_drift_2 + directed_forgetting.proactive_interference_EZ_drift_2 + dot_pattern_expectancy.AY.BY_EZ_drift_2 + dot_pattern_expectancy.BX.BY_EZ_drift_2 + dot_pattern_expectancy.EZ_drift_2 + local_global_letter.EZ_drift_2 + local_global_letter.conflict_EZ_drift_2 + local_global_letter.global_bias_EZ_drift_2 + local_global_letter.switch_cost_EZ_drift_2 + motor_selective_stop_signal.EZ_drift_2 + recent_probes.EZ_drift_2 + recent_probes.proactive_interference_EZ_drift_2 + shape_matching.EZ_drift_2 + shape_matching.stimulus_interference_EZ_drift_2 + simon.EZ_drift_2 + simon.simon_EZ_drift_2 + stim_selective_stop_signal.EZ_drift_2 + stop_signal.EZ_drift_2 + stroop.EZ_drift_2 + stroop.stroop_EZ_drift_2 + threebytwo.EZ_drift_2 + threebytwo.cue_switch_cost_EZ_drift_100.0_2 + threebytwo.cue_switch_cost_EZ_drift_900.0_2 + threebytwo.task_switch_cost_EZ_drift_100.0_2 + threebytwo.task_switch_cost_EZ_drift_900.0_2
thresh_2 =~ adaptive_n_back.EZ_thresh_2 + attention_network_task.EZ_thresh_2 + attention_network_task.alerting_EZ_thresh_2 + attention_network_task.conflict_EZ_thresh_2 + attention_network_task.orienting_EZ_thresh_2 + choice_reaction_time.EZ_thresh_2 + directed_forgetting.EZ_thresh_2 + directed_forgetting.proactive_interference_EZ_thresh_2 + dot_pattern_expectancy.AY.BY_EZ_thresh_2 + dot_pattern_expectancy.BX.BY_EZ_thresh_2 + dot_pattern_expectancy.EZ_thresh_2 + local_global_letter.EZ_thresh_2 + local_global_letter.conflict_EZ_thresh_2 + local_global_letter.global_bias_EZ_thresh_2 + local_global_letter.switch_cost_EZ_thresh_2 + motor_selective_stop_signal.EZ_thresh_2 + recent_probes.EZ_thresh_2 + recent_probes.proactive_interference_EZ_thresh_2 + shape_matching.EZ_thresh_2 + shape_matching.stimulus_interference_EZ_thresh_2 + simon.EZ_thresh_2 + simon.simon_EZ_thresh_2 + stim_selective_stop_signal.EZ_thresh_2 + stop_signal.EZ_thresh_2 + stroop.EZ_thresh_2 + stroop.stroop_EZ_thresh_2 + threebytwo.EZ_thresh_2 + threebytwo.cue_switch_cost_EZ_thresh_100.0_2 + threebytwo.cue_switch_cost_EZ_thresh_900.0_2 + threebytwo.task_switch_cost_EZ_thresh_100.0_2 + threebytwo.task_switch_cost_EZ_thresh_900.0_2
non_dec_2 =~ adaptive_n_back.EZ_non_decision_2 + attention_network_task.EZ_non_decision_2 + attention_network_task.alerting_EZ_non_decision_2 + attention_network_task.conflict_EZ_non_decision_2 + attention_network_task.orienting_EZ_non_decision_2 + choice_reaction_time.EZ_non_decision_2 + directed_forgetting.EZ_non_decision_2 + directed_forgetting.proactive_interference_EZ_non_decision_2 + dot_pattern_expectancy.AY.BY_EZ_non_decision_2 + dot_pattern_expectancy.BX.BY_EZ_non_decision_2 + dot_pattern_expectancy.EZ_non_decision_2 + local_global_letter.EZ_non_decision_2 + local_global_letter.switch_cost_EZ_non_decision_2 + motor_selective_stop_signal.EZ_non_decision_2 + recent_probes.EZ_non_decision_2 + recent_probes.proactive_interference_EZ_non_decision_2 + shape_matching.EZ_non_decision_2 + shape_matching.stimulus_interference_EZ_non_decision_2 + simon.EZ_non_decision_2 + simon.simon_EZ_non_decision_2 + stim_selective_stop_signal.EZ_non_decision_2 + stop_signal.EZ_non_decision_2 + stroop.EZ_non_decision_2 + stroop.stroop_EZ_non_decision_2 + threebytwo.EZ_non_decision_2 + threebytwo.cue_switch_cost_EZ_non_decision_100.0_2 + threebytwo.cue_switch_cost_EZ_non_decision_900.0_2 + threebytwo.task_switch_cost_EZ_non_decision_100.0_2 + threebytwo.task_switch_cost_EZ_non_decision_900.0_2

# intercepts

# residual covariances
adaptive_n_back.EZ_drift_1 ~~ adaptive_n_back.EZ_drift_2
 attention_network_task.EZ_drift_1 ~~ attention_network_task.EZ_drift_2
 attention_network_task.alerting_EZ_drift_1 ~~ attention_network_task.alerting_EZ_drift_2
 attention_network_task.conflict_EZ_drift_1 ~~ attention_network_task.conflict_EZ_drift_2
 attention_network_task.orienting_EZ_drift_1 ~~ attention_network_task.orienting_EZ_drift_2
 choice_reaction_time.EZ_drift_1 ~~ choice_reaction_time.EZ_drift_2
 directed_forgetting.EZ_drift_1 ~~ directed_forgetting.EZ_drift_2
 directed_forgetting.proactive_interference_EZ_drift_1 ~~ directed_forgetting.proactive_interference_EZ_drift_2
 dot_pattern_expectancy.AY.BY_EZ_drift_1 ~~ dot_pattern_expectancy.AY.BY_EZ_drift_2
 dot_pattern_expectancy.BX.BY_EZ_drift_1 ~~ dot_pattern_expectancy.BX.BY_EZ_drift_2
 dot_pattern_expectancy.EZ_drift_1 ~~ dot_pattern_expectancy.EZ_drift_2
 local_global_letter.EZ_drift_1 ~~ local_global_letter.EZ_drift_2
 local_global_letter.conflict_EZ_drift_1 ~~ local_global_letter.conflict_EZ_drift_2
 local_global_letter.global_bias_EZ_drift_1 ~~ local_global_letter.global_bias_EZ_drift_2
 local_global_letter.switch_cost_EZ_drift_1 ~~ local_global_letter.switch_cost_EZ_drift_2
 motor_selective_stop_signal.EZ_drift_1 ~~ motor_selective_stop_signal.EZ_drift_2
 recent_probes.EZ_drift_1 ~~ recent_probes.EZ_drift_2
 recent_probes.proactive_interference_EZ_drift_1 ~~ recent_probes.proactive_interference_EZ_drift_2
 shape_matching.EZ_drift_1 ~~ shape_matching.EZ_drift_2
 shape_matching.stimulus_interference_EZ_drift_1 ~~ shape_matching.stimulus_interference_EZ_drift_2
 simon.EZ_drift_1 ~~ simon.EZ_drift_2
 simon.simon_EZ_drift_1 ~~ simon.simon_EZ_drift_2
 stim_selective_stop_signal.EZ_drift_1 ~~ stim_selective_stop_signal.EZ_drift_2
 stop_signal.EZ_drift_1 ~~ stop_signal.EZ_drift_2
 stroop.EZ_drift_1 ~~ stroop.EZ_drift_2
 stroop.stroop_EZ_drift_1 ~~ stroop.stroop_EZ_drift_2
 threebytwo.EZ_drift_1 ~~ threebytwo.EZ_drift_2
 threebytwo.cue_switch_cost_EZ_drift_100.0_1 ~~ threebytwo.cue_switch_cost_EZ_drift_100.0_2
 threebytwo.cue_switch_cost_EZ_drift_900.0_1 ~~ threebytwo.cue_switch_cost_EZ_drift_900.0_2
 threebytwo.task_switch_cost_EZ_drift_100.0_1 ~~ threebytwo.task_switch_cost_EZ_drift_100.0_2
 threebytwo.task_switch_cost_EZ_drift_900.0_1 ~~ threebytwo.task_switch_cost_EZ_drift_900.0_2
adaptive_n_back.EZ_thresh_1 ~~ adaptive_n_back.EZ_thresh_2
 attention_network_task.EZ_thresh_1 ~~ attention_network_task.EZ_thresh_2
 attention_network_task.alerting_EZ_thresh_1 ~~ attention_network_task.alerting_EZ_thresh_2
 attention_network_task.conflict_EZ_thresh_1 ~~ attention_network_task.conflict_EZ_thresh_2
 attention_network_task.orienting_EZ_thresh_1 ~~ attention_network_task.orienting_EZ_thresh_2
 choice_reaction_time.EZ_thresh_1 ~~ choice_reaction_time.EZ_thresh_2
 directed_forgetting.EZ_thresh_1 ~~ directed_forgetting.EZ_thresh_2
 directed_forgetting.proactive_interference_EZ_thresh_1 ~~ directed_forgetting.proactive_interference_EZ_thresh_2
 dot_pattern_expectancy.AY.BY_EZ_thresh_1 ~~ dot_pattern_expectancy.AY.BY_EZ_thresh_2
 dot_pattern_expectancy.BX.BY_EZ_thresh_1 ~~ dot_pattern_expectancy.BX.BY_EZ_thresh_2
 dot_pattern_expectancy.EZ_thresh_1 ~~ dot_pattern_expectancy.EZ_thresh_2
 local_global_letter.EZ_thresh_1 ~~ local_global_letter.EZ_thresh_2
 local_global_letter.conflict_EZ_thresh_1 ~~ local_global_letter.conflict_EZ_thresh_2
 local_global_letter.global_bias_EZ_thresh_1 ~~ local_global_letter.global_bias_EZ_thresh_2
 local_global_letter.switch_cost_EZ_thresh_1 ~~ local_global_letter.switch_cost_EZ_thresh_2
 motor_selective_stop_signal.EZ_thresh_1 ~~ motor_selective_stop_signal.EZ_thresh_2
 recent_probes.EZ_thresh_1 ~~ recent_probes.EZ_thresh_2
 recent_probes.proactive_interference_EZ_thresh_1 ~~ recent_probes.proactive_interference_EZ_thresh_2
 shape_matching.EZ_thresh_1 ~~ shape_matching.EZ_thresh_2
 shape_matching.stimulus_interference_EZ_thresh_1 ~~ shape_matching.stimulus_interference_EZ_thresh_2
 simon.EZ_thresh_1 ~~ simon.EZ_thresh_2
 simon.simon_EZ_thresh_1 ~~ simon.simon_EZ_thresh_2
 stim_selective_stop_signal.EZ_thresh_1 ~~ stim_selective_stop_signal.EZ_thresh_2
 stop_signal.EZ_thresh_1 ~~ stop_signal.EZ_thresh_2
 stroop.EZ_thresh_1 ~~ stroop.EZ_thresh_2
 stroop.stroop_EZ_thresh_1 ~~ stroop.stroop_EZ_thresh_2
 threebytwo.EZ_thresh_1 ~~ threebytwo.EZ_thresh_2
 threebytwo.cue_switch_cost_EZ_thresh_100.0_1 ~~ threebytwo.cue_switch_cost_EZ_thresh_100.0_2
 threebytwo.cue_switch_cost_EZ_thresh_900.0_1 ~~ threebytwo.cue_switch_cost_EZ_thresh_900.0_2
 threebytwo.task_switch_cost_EZ_thresh_100.0_1 ~~ threebytwo.task_switch_cost_EZ_thresh_100.0_2
 threebytwo.task_switch_cost_EZ_thresh_900.0_1 ~~ threebytwo.task_switch_cost_EZ_thresh_900.0_2
adaptive_n_back.EZ_non_decision_1 ~~ adaptive_n_back.EZ_non_decision_2
 attention_network_task.EZ_non_decision_1 ~~ attention_network_task.EZ_non_decision_2
 attention_network_task.alerting_EZ_non_decision_1 ~~ attention_network_task.alerting_EZ_non_decision_2
 attention_network_task.conflict_EZ_non_decision_1 ~~ attention_network_task.conflict_EZ_non_decision_2
 attention_network_task.orienting_EZ_non_decision_1 ~~ attention_network_task.orienting_EZ_non_decision_2
 choice_reaction_time.EZ_non_decision_1 ~~ choice_reaction_time.EZ_non_decision_2
 directed_forgetting.EZ_non_decision_1 ~~ directed_forgetting.EZ_non_decision_2
 directed_forgetting.proactive_interference_EZ_non_decision_1 ~~ directed_forgetting.proactive_interference_EZ_non_decision_2
 dot_pattern_expectancy.AY.BY_EZ_non_decision_1 ~~ dot_pattern_expectancy.AY.BY_EZ_non_decision_2
 dot_pattern_expectancy.BX.BY_EZ_non_decision_1 ~~ dot_pattern_expectancy.BX.BY_EZ_non_decision_2
 dot_pattern_expectancy.EZ_non_decision_1 ~~ dot_pattern_expectancy.EZ_non_decision_2
 local_global_letter.EZ_non_decision_1 ~~ local_global_letter.EZ_non_decision_2
 local_global_letter.switch_cost_EZ_non_decision_1 ~~ local_global_letter.switch_cost_EZ_non_decision_2
 motor_selective_stop_signal.EZ_non_decision_1 ~~ motor_selective_stop_signal.EZ_non_decision_2
 recent_probes.EZ_non_decision_1 ~~ recent_probes.EZ_non_decision_2
 recent_probes.proactive_interference_EZ_non_decision_1 ~~ recent_probes.proactive_interference_EZ_non_decision_2
 shape_matching.EZ_non_decision_1 ~~ shape_matching.EZ_non_decision_2
 shape_matching.stimulus_interference_EZ_non_decision_1 ~~ shape_matching.stimulus_interference_EZ_non_decision_2
 simon.EZ_non_decision_1 ~~ simon.EZ_non_decision_2
 simon.simon_EZ_non_decision_1 ~~ simon.simon_EZ_non_decision_2
 stim_selective_stop_signal.EZ_non_decision_1 ~~ stim_selective_stop_signal.EZ_non_decision_2
 stop_signal.EZ_non_decision_1 ~~ stop_signal.EZ_non_decision_2
 stroop.EZ_non_decision_1 ~~ stroop.EZ_non_decision_2
 stroop.stroop_EZ_non_decision_1 ~~ stroop.stroop_EZ_non_decision_2
 threebytwo.EZ_non_decision_1 ~~ threebytwo.EZ_non_decision_2
 threebytwo.cue_switch_cost_EZ_non_decision_100.0_1 ~~ threebytwo.cue_switch_cost_EZ_non_decision_100.0_2
 threebytwo.cue_switch_cost_EZ_non_decision_900.0_1 ~~ threebytwo.cue_switch_cost_EZ_non_decision_900.0_2
 threebytwo.task_switch_cost_EZ_non_decision_100.0_1 ~~ threebytwo.task_switch_cost_EZ_non_decision_100.0_2
 threebytwo.task_switch_cost_EZ_non_decision_900.0_1 ~~ threebytwo.task_switch_cost_EZ_non_decision_900.0_2

#latent means
drift_1 ~ 0*1
drift_2 ~ 1
thresh_1 ~ 0*1
thresh_2 ~ 1
non_dec_1 ~ 0*1
non_dec_2 ~ 1
'
```

```{r}
ez.latent.time.model <- paste0("\ndrift_1=~", paste0(drift_vars_sub, "_1", collapse= " + "),
       "\nthresh_1=~", paste0(thresh_vars_sub, "_1", collapse= " + "),
       "\nnon_dec_1=~", paste0(non_dec_vars_sub, "_1", collapse= " + "), 
       "\ndrift_2=~", paste0(drift_vars_sub, "_2", collapse= " + "),
       "\nthresh_2=~", paste0(thresh_vars_sub, "_2", collapse= " + "),
       "\nnon_dec_2=~", paste0(non_dec_vars_sub, "_2", collapse= " + "), 
       paste0("\n",drift_vars_sub, "_1"," ~~ ", drift_vars_sub, "_2", collapse = "\n"),
       paste0("\n",thresh_vars_sub, "_1"," ~~ ", thresh_vars_sub, "_2", collapse = "\n"),
       paste0("\n",non_dec_vars_sub, "_1"," ~~ ", non_dec_vars_sub, "_2", collapse = "\n"),
       paste0('\ndrift_1 ~ 0*1', '\ndrift_2 ~ 1','\nthresh_1 ~ 0*1','\nthresh_2 ~ 1','\nnon_dec_1 ~ 0*1','\nnon_dec_2 ~ 1',collapse = "\n"),
        collapse = "\n ")
```


```{r eval=FALSE}
ez.latent.time.model.fit <- cfa(ez.latent.time.model, sample.cov = COV, sample.mean = MEAN, sample.nobs = 150)
```


```{r eval=FALSE}
summary(ez.latent.time.model.fit, fit.measures = TRUE, standardized = TRUE)
```
