---
title: 'SRO DDM Prediction comparison'
output:
github_document:
toc: yes
toc_float: yes
---

Trying to get to the bottom of the differences between my and Ian's prediction analyses.

Problem: In his analyses Ian found very low cross-validated $R^2$ values when predicting demographic factors using the task ontology. In my analyses using factors of EZ DDM variables and raw RT's and accuracies I found relationships that were >0.

Approach: Our analyses differ in multiple aspects both in the predicted data (DV data) and predictor data (IV data) so we checked each of them to find where the difference may lay.  

#1. DV data  
We both calculated 9 factor solutions for the demographic items separately. our factor scores are essentially the same as can be seen in the last plot [here](https://zenkavi.github.io/SRO_DDM_Analyses/output/reports/DemographicFactors.nb.html).  

Still, checked whether there were any differences predicting his or my factor scores. 

The plot below suggests  
- There are no systematic differences depending on the DV data  
- The largest difference is between predicting using n=522 or n=150  
- Raw measures seem to have some >0 association too  

```{r}
helper_func_path = 'https://raw.githubusercontent.com/zenkavi/SRO_Retest_Analyses/master/code/helper_functions/'

eval(parse(text = getURL(paste0(helper_func_path,'sem.R'), ssl.verifypeer = FALSE)))

ddm_workspace_scripts = 'https://raw.githubusercontent.com/zenkavi/SRO_DDM_Analyses/master/code/workspace_scripts/'

eval(parse(text = getURL(paste0(ddm_workspace_scripts,'ddm_measure_labels.R'), ssl.verifypeer = FALSE)))

input_path='/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_DDM_Analyses/input/'

out_device = "jpeg"

fig_path = '/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_DDM_Analyses/output/figures/'
```

```{r warning=FALSE, message=FALSE}
iv_dfs = c("ez_t1_522_fa_3_scores", "ez_t1_fa_3_scores", "res_clean_test_data_raw", "res_clean_test_data_ez")
dv_dfs = c('ian_demog_scores', 'demog_fa_scores_t1')

out = data.frame()

for(iv_data in iv_dfs){
  for(dv_data in dv_dfs){
    file_name = paste0(input_path,'prediction/pred_out_', iv_data,'_', dv_data, '.csv')
    
    d = read.csv(file_name) 
    
    if(iv_data %in% c("ez_t1_522_fa_3_scores", "ez_t1_fa_3_scores")){
      d = d %>%
        mutate(RsquaredSE = RsquaredSD/sqrt(10),
         iv=as.character(iv),
         dv = as.character(dv),
         iv_data = as.character(iv_data),
         dv_data = as.character(dv_data)) %>%
  select(dv, iv, Rsquared, RsquaredSE, iv_data, dv_data)
    }
    
    if(iv_data =="res_clean_test_data_raw"){
      d = d %>%
        mutate(iv=as.character(iv),
         iv=gsub(".ReflogTr", "",iv),
         iv=gsub(".logTr", "",iv)) %>% 
  left_join(measure_labels %>% select(dv, rt_acc) %>% rename(iv=dv), by="iv") %>%
  group_by(dv, rt_acc) %>%
  summarise(RsquaredSE = sem(Rsquared),
            Rsquared = mean(Rsquared),
            iv_data = as.character(unique(iv_data)),
            dv_data = as.character(unique(dv_data))) %>%
  rename(iv = rt_acc) %>%
  ungroup() %>%
  mutate(dv = as.character(dv)) %>%
  select(dv, iv, Rsquared, RsquaredSE, iv_data, dv_data)
    }
    
    if(iv_data == "res_clean_test_data_ez"){
      d = d%>%
        mutate(par = ifelse(grepl("drift",iv), "drift_rate", ifelse(grepl("thresh", iv), "threshold", ifelse(grepl("non_decision", iv), "non_decision", NA)))) %>% 
  group_by(dv, par) %>%
  summarise(RsquaredSE = sem(Rsquared),
            Rsquared = mean(Rsquared),
            iv_data = as.character(unique(iv_data)),
            dv_data = as.character(unique(dv_data))) %>%
  rename(iv = par) %>%
  ungroup() %>%
  mutate(dv = as.character(dv)) %>%
  select(dv, iv, Rsquared, RsquaredSE, iv_data, dv_data)
    }
    out = out %>%
      bind_rows(d)
  }
}
```

```{r}
p = out %>% 
  mutate(iv = factor(iv, levels = c("drift_rate", "threshold", "non_decision","accuracy", "rt"), labels=c("drift rate", "threshold", "non-decision","accuracy", "rt")),
         dv = factor(dv, levels = c('Drug_Use','Mental_Health','Problem_Drinking','Daily_Smoking','Binge_Drinking','Obesity','Lifetime_Smoking','Unsafe_Drinking','Income_LifeMilestones'), labels = c('Drug Use','Mental Health','Problem Drinking','Daily Smoking','Binge Drinking','Obesity','Lifetime Smoking','Unsafe Drinking','Income/Life Milestones')),
         iv_data = factor(iv_data, levels = c("ez_t1_522_fa_3_scores", "ez_t1_fa_3_scores", "res_clean_test_data_ez", "res_clean_test_data_raw"),
                          labels = c("EZ factors (n=522)", "EZ factors (n=150)", "EZ measures", "Raw measures"))) %>%
  ggplot(aes(iv, Rsquared, fill=iv, alpha=dv_data))+
  geom_bar(stat="identity", position = position_dodge())+
  geom_errorbar(aes(ymin=Rsquared-RsquaredSE, ymax=Rsquared+RsquaredSE, color=iv), position=position_dodge(width=0.9), width=0.1)+
  facet_grid(iv_data~dv, scales='free_x', labeller = label_wrap_gen(width = 2, multi_line = TRUE))+
  theme(legend.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.text = element_text(size=16),
        strip.text = element_text(size=16),
        axis.title.y = element_text(size=16),
        axis.text.y= element_text(size=14))+
  xlab("")+
  ylab(expression(R^{2}))+
  scale_alpha_manual(values = c(0.5, 1),
                     breaks = c("demog_fa_scores_t1", "ian_demog_scores"),
                     labels = c("My demog factors", "Ian demog factors"))
  # guides(alpha=FALSE)

ggsave(paste0('pred_dv_data_comparison.', out_device), plot=p, device = out_device, path = fig_path, width = 20, height = 9, units = "in")
```

Larger figure [here](https://github.com/zenkavi/SRO_DDM_Analyses/blob/master/output/figures/pred_dv_data_comparison.jpeg)

```{r}
fig_name = 'pred_dv_data_comparison.jpeg'

knitr::include_graphics(paste0(fig_path, fig_name))
```

#2. IV data

Based on the above plot I wondered:  
a. Is there something weird about retest sample (n=150)  
b. If the factor model for the EZ variables is not good due to linear dependencies between the measures going into the model do the results change if we fit the model on a subset of the EZ variables (specifically condition variables only)  

Thus the IV data can vary on a few dimensions:  

- Sample the factor models are based on: The models depicted below are always fit on the whole sample 
- Sample the prediction is done on: Even though all models are fit on the whole sample for prediction I used either all the scores or smaller subsets (retest sample and non-retest sample)
- Variables that go in to the factor model: Modeling either all EZ variables or just EZ condition variables

In the plot n=150 refers to the retest sample, n=372 subjects who aren't in the retest sample (non-retest sample)


```{r warning=FALSE, message=FALSE}
iv_dfs = c('ez_t1_522_fa_3_scores', 'ez_t1_522_fa_3_scores_t2subs', 'ez_t1_522_fa_3_scores_nont2subs', 'ez_t1_522_fa_3_condition_scores', 'ez_t1_522_fa_3_condition_scores_t2subs', 'ez_t1_522_fa_3_condition_scores_nont2subs', 'res_clean_test_data_ez','res_clean_test_data_ez_522','res_clean_test_data_ez_nont2subs','res_clean_test_data_raw','res_clean_test_data_raw_522','res_clean_test_data_raw_nont2subs' )

out = data.frame()

for(iv_data in iv_dfs){
  for(dv_data in dv_dfs){
    file_name = paste0(input_path,'prediction/pred_out_', iv_data,'_', dv_data, '.csv')
    
    d = read.csv(file_name) 
    
    if(iv_data %in% c('ez_t1_522_fa_3_scores', 'ez_t1_522_fa_3_scores_t2subs', 'ez_t1_522_fa_3_scores_nont2subs', 'ez_t1_522_fa_3_condition_scores', 'ez_t1_522_fa_3_condition_scores_t2subs', 'ez_t1_522_fa_3_condition_scores_nont2subs')){
      d = d %>%
        mutate(RsquaredSE = RsquaredSD/sqrt(10),
         iv=as.character(iv),
         dv = as.character(dv),
         iv_data = as.character(iv_data),
         dv_data = as.character(dv_data)) %>%
  select(dv, iv, Rsquared, RsquaredSE, iv_data, dv_data)
    }
    
    if(iv_data %in% c("res_clean_test_data_raw_522", "res_clean_test_data_raw", "res_clean_test_data_raw_nont2subs")){
      d = d %>%
        mutate(iv=as.character(iv),
         iv=gsub(".ReflogTr", "",iv),
         iv=gsub(".logTr", "",iv)) %>% 
  left_join(measure_labels %>% select(dv, rt_acc) %>% rename(iv=dv), by="iv") %>%
  group_by(dv, rt_acc) %>%
  summarise(RsquaredSE = sem(Rsquared),
            Rsquared = mean(Rsquared),
            iv_data = as.character(unique(iv_data)),
            dv_data = as.character(unique(dv_data))) %>%
  rename(iv = rt_acc) %>%
  ungroup() %>%
  mutate(dv = as.character(dv)) %>%
  select(dv, iv, Rsquared, RsquaredSE, iv_data, dv_data)
    }
    
    if(iv_data %in% c("res_clean_test_data_ez_522", "res_clean_test_data_ez", "res_clean_test_data_ez_nont2subs")){
      d = d%>%
        mutate(par = ifelse(grepl("drift",iv), "drift_rate", ifelse(grepl("thresh", iv), "threshold", ifelse(grepl("non_decision", iv), "non_decision", NA)))) %>% 
  group_by(dv, par) %>%
  summarise(RsquaredSE = sem(Rsquared),
            Rsquared = mean(Rsquared),
            iv_data = as.character(unique(iv_data)),
            dv_data = as.character(unique(dv_data))) %>%
  rename(iv = par) %>%
  ungroup() %>%
  mutate(dv = as.character(dv)) %>%
  select(dv, iv, Rsquared, RsquaredSE, iv_data, dv_data)
    }
    out = out %>%
      bind_rows(d)
  }
}
```

```{r}
p = out %>% 
  mutate(iv = factor(iv, levels = c("drift_rate", "threshold", "non_decision", "accuracy", "rt"), labels=c("drift rate", "threshold", "non-decision","accuracy", "rt")),
         dv = factor(dv, levels = c('Drug_Use','Mental_Health','Problem_Drinking','Daily_Smoking','Binge_Drinking','Obesity','Lifetime_Smoking','Unsafe_Drinking','Income_LifeMilestones'), labels = c('Drug Use','Mental Health','Problem Drinking','Daily Smoking','Binge Drinking','Obesity','Lifetime Smoking','Unsafe Drinking','Income/Life Milestones')),
         iv_data_2 = ifelse(iv_data %in% c('ez_t1_522_fa_3_scores', 'ez_t1_522_fa_3_scores_t2subs', 'ez_t1_522_fa_3_scores_nont2subs'),"EZ all var factors",ifelse(iv_data %in% c('ez_t1_522_fa_3_condition_scores', 'ez_t1_522_fa_3_condition_scores_t2subs', 'ez_t1_522_fa_3_condition_scores_nont2subs'),"EZ cond var factors", ifelse(iv_data %in% c('res_clean_test_data_ez_522', 'res_clean_test_data_ez', 'res_clean_test_data_ez_nont2subs'), "EZ measures", ifelse(iv_data %in% c('res_clean_test_data_raw_522','res_clean_test_data_raw','res_clean_test_data_raw_nont2subs'),"Raw measures",NA)))),
         sample = ifelse(iv_data %in% c('ez_t1_522_fa_3_scores','ez_t1_522_fa_3_condition_scores', 'res_clean_test_data_ez_522','res_clean_test_data_raw_522'),522,ifelse(iv_data %in% c('ez_t1_522_fa_3_scores_nont2subs', 'ez_t1_522_fa_3_condition_scores_nont2subs', 'res_clean_test_data_ez_nont2subs', 'res_clean_test_data_raw_nont2subs'),372, ifelse(iv_data %in% c('ez_t1_522_fa_3_scores_t2subs', 'ez_t1_522_fa_3_condition_scores_t2subs', 'res_clean_test_data_ez', 'res_clean_test_data_raw'),150, NA))),
         sample = factor(sample, levels = c(150, 372, 522), labels = c(150, 372, 522))) %>%
  ggplot(aes(iv, Rsquared, fill=iv, alpha=sample))+
  geom_bar(stat="identity", position = position_dodge())+
  geom_errorbar(aes(ymin=Rsquared-RsquaredSE, ymax=Rsquared+RsquaredSE, color=iv), position=position_dodge(width=0.9), width=0.1)+
  facet_grid(iv_data_2~dv, scales='free_x', labeller = label_wrap_gen(width = 2, multi_line = TRUE))+
  theme(legend.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.text = element_text(size=16),
        strip.text = element_text(size=16),
        axis.title.y = element_text(size=16),
        axis.text.y= element_text(size=14))+
  xlab("")+
  ylab(expression(R^{2}))+
  scale_alpha_manual(values = c(0.33,0.66, 1))

ggsave(paste0('pred_iv_data_comparison.', out_device), plot=p, device = out_device, path = fig_path, width = 20, height = 9, units = "in")
```

Larger figure [here](https://github.com/zenkavi/SRO_DDM_Analyses/blob/master/output/figures/pred_iv_data_comparison.jpeg)

```{r}
fig_name = 'pred_iv_data_comparison.jpeg'

knitr::include_graphics(paste0(fig_path, fig_name))
```

## RMSE instead of $R^2$

```{r}
iv_dfs = c('ez_t1_522_fa_3_condition_scores', 'ez_t1_522_fa_3_condition_scores_t2subs', 'ez_t1_522_fa_3_condition_scores_nont2subs')
dv_dfs = c('demog_fa_scores_t1')

out = data.frame()

for(iv_data in iv_dfs){
  for(dv_data in dv_dfs){
    file_name = paste0(input_path,'prediction/pred_out_', iv_data,'_', dv_data, '.csv')
    
    d = read.csv(file_name) 
    
    if(iv_data %in% c('ez_t1_522_fa_3_scores', 'ez_t1_522_fa_3_scores_t2subs', 'ez_t1_522_fa_3_scores_nont2subs', 'ez_t1_522_fa_3_condition_scores', 'ez_t1_522_fa_3_condition_scores_t2subs', 'ez_t1_522_fa_3_condition_scores_nont2subs')){
      d = d %>%
        mutate(RMSESE = RMSESD/sqrt(10),
         iv=as.character(iv),
         dv = as.character(dv),
         iv_data = as.character(iv_data),
         dv_data = as.character(dv_data)) %>%
  select(dv, iv, RMSE, RMSESE, iv_data, dv_data)
    }
    
    if(iv_data %in% c("res_clean_test_data_raw_522", "res_clean_test_data_raw", "res_clean_test_data_raw_nont2subs")){
      d = d %>%
        mutate(iv=as.character(iv),
         iv=gsub(".ReflogTr", "",iv),
         iv=gsub(".logTr", "",iv)) %>% 
  left_join(measure_labels %>% select(dv, rt_acc) %>% rename(iv=dv), by="iv") %>%
  group_by(dv, rt_acc) %>%
  summarise(RMSESE = sem(RMSE),
            RMSE = mean(RMSE),
            iv_data = as.character(unique(iv_data)),
            dv_data = as.character(unique(dv_data))) %>%
  rename(iv = rt_acc) %>%
  ungroup() %>%
  mutate(dv = as.character(dv)) %>%
  select(dv, iv, RMSE, RMSESE, iv_data, dv_data)
    }
    
    if(iv_data %in% c("res_clean_test_data_ez_522", "res_clean_test_data_ez", "res_clean_test_data_ez_nont2subs")){
      d = d%>%
        mutate(par = ifelse(grepl("drift",iv), "drift_rate", ifelse(grepl("thresh", iv), "threshold", ifelse(grepl("non_decision", iv), "non_decision", NA)))) %>% 
  group_by(dv, par) %>%
  summarise(RMSESE = sem(RMSE),
            RMSE = mean(RMSE),
            iv_data = as.character(unique(iv_data)),
            dv_data = as.character(unique(dv_data))) %>%
  rename(iv = par) %>%
  ungroup() %>%
  mutate(dv = as.character(dv)) %>%
  select(dv, iv, RMSE, RMSESE, iv_data, dv_data)
    }
    out = out %>%
      bind_rows(d)
  }
}
```


```{r}
out %>% 
  mutate(iv = factor(iv, levels = c("drift_rate", "threshold", "non_decision"), labels=c("drift rate", "threshold", "non-decision")),
         dv = factor(dv, levels = c('Drug_Use','Mental_Health','Problem_Drinking','Daily_Smoking','Binge_Drinking','Obesity','Lifetime_Smoking','Unsafe_Drinking','Income_LifeMilestones'), labels = c('Drug Use','Mental Health','Problem Drinking','Daily Smoking','Binge Drinking','Obesity','Lifetime Smoking','Unsafe Drinking','Income/Life Milestones')),
         iv_data_2 = ifelse(iv_data %in% c('ez_t1_522_fa_3_scores', 'ez_t1_522_fa_3_scores_t2subs', 'ez_t1_522_fa_3_scores_nont2subs'),"EZ all var factors",ifelse(iv_data %in% c('ez_t1_522_fa_3_condition_scores', 'ez_t1_522_fa_3_condition_scores_t2subs', 'ez_t1_522_fa_3_condition_scores_nont2subs'),"EZ cond var factors", ifelse(iv_data %in% c('res_clean_test_data_ez_522', 'res_clean_test_data_ez', 'res_clean_test_data_ez_nont2subs'), "EZ measures", ifelse(iv_data %in% c('res_clean_test_data_raw_522','res_clean_test_data_raw','res_clean_test_data_raw_nont2subs'),"Raw measures",NA)))),
         sample = ifelse(iv_data %in% c('ez_t1_522_fa_3_scores','ez_t1_522_fa_3_condition_scores', 'res_clean_test_data_ez_522','res_clean_test_data_raw_522'),522,ifelse(iv_data %in% c('ez_t1_522_fa_3_scores_nont2subs', 'ez_t1_522_fa_3_condition_scores_nont2subs', 'res_clean_test_data_ez_nont2subs', 'res_clean_test_data_raw_nont2subs'),372, ifelse(iv_data %in% c('ez_t1_522_fa_3_scores_t2subs', 'ez_t1_522_fa_3_condition_scores_t2subs', 'res_clean_test_data_ez', 'res_clean_test_data_raw'),150, NA))),
         sample = factor(sample, levels = c(150, 372, 522), labels = c(150, 372, 522))) %>%
  ggplot(aes(iv, RMSE, fill=iv, alpha=sample))+
  geom_bar(stat="identity", position = position_dodge())+
  geom_errorbar(aes(ymin=RMSE-RMSESE, ymax=RMSE+RMSESE, color=iv), position=position_dodge(width=0.9), width=0.1)+
  facet_grid(.~dv, scales='free_x', labeller = label_wrap_gen(width = 2, multi_line = TRUE))+
  theme(legend.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.text = element_text(size=16),
        strip.text = element_text(size=16),
        axis.title.y = element_text(size=16),
        axis.text.y= element_text(size=14))+
  xlab("")+
  ylab("RMSE")+
  scale_alpha_manual(values = c(0.33,0.66, 1))

ggsave(paste0('pred_iv_data_comparison_RMSE.', out_device), device = out_device, path = fig_path, width = 20, height = 9, units = "in")
```

# Correlations across folds 

```{r}
iv_dfs = c("ez_t1_522_fa_3_condition_scores", "ez_t1_522_fa_3_condition_scores_t2subs", "ez_t1_522_fa_3_condition_scores_nont2subs")
dv_dfs = c('demog_fa_scores_t1')

out = data.frame()

for(iv_data in iv_dfs){
  for(dv_data in dv_dfs){
    file_name = paste0(input_path,'prediction/pred_fold_cors_', iv_data,'_', dv_data, '.csv')
    
    d = read.csv(file_name) 
    
    out = out %>%
      bind_rows(d)
  }
}
```

```{r}
out_m = out %>%
  group_by(dv, iv,iv_data) %>%
  summarise(m_R = mean(R),
            sem_R = sem(R))

out = out %>%
  mutate(iv = factor(iv, levels = c("drift_rate", "threshold", "non_decision"), labels=c("drift rate", "threshold", "non-decision")),
         dv = factor(dv, levels = c('Drug_Use','Mental_Health','Problem_Drinking','Daily_Smoking','Binge_Drinking','Obesity','Lifetime_Smoking','Unsafe_Drinking','Income_LifeMilestones'), labels = c('Drug Use','Mental Health','Problem Drinking','Daily Smoking','Binge Drinking','Obesity','Lifetime Smoking','Unsafe Drinking','Income/Life Milestones')),
         iv_data = factor(iv_data, levels = c("ez_t1_522_fa_3_condition_scores_t2subs", "ez_t1_522_fa_3_condition_scores_nont2subs", "ez_t1_522_fa_3_condition_scores"), labels = c(150, 372, 522)))


```


```{r}
out_m %>%
  ungroup() %>%
  mutate(iv = factor(iv, levels = c("drift_rate", "threshold", "non_decision"), labels=c("drift rate", "threshold", "non-decision")),
         dv = factor(dv, levels = c('Drug_Use','Mental_Health','Problem_Drinking','Daily_Smoking','Binge_Drinking','Obesity','Lifetime_Smoking','Unsafe_Drinking','Income_LifeMilestones'), labels = c('Drug Use','Mental Health','Problem Drinking','Daily Smoking','Binge Drinking','Obesity','Lifetime Smoking','Unsafe Drinking','Income/Life Milestones')),
         iv_data = factor(iv_data, levels = c("ez_t1_522_fa_3_condition_scores_t2subs", "ez_t1_522_fa_3_condition_scores_nont2subs", "ez_t1_522_fa_3_condition_scores"), labels = c("150","372","522"))) %>%
  ggplot(aes(iv, m_R))+
  geom_bar(stat="identity", position=position_dodge(width=0.9), aes(alpha=iv_data, fill=iv))+
  geom_errorbar(aes(ymin=m_R-sem_R, ymax=m_R+sem_R, col=iv, alpha=iv_data), position = position_dodge(width = 0.9), width=0.1)+
  geom_point(data=out, aes(iv, R, alpha=iv_data, color=iv), position = position_dodge(width=0.9))+
  facet_grid(~dv,labeller = label_wrap_gen(width = 2, multi_line = TRUE))+
  xlab("")+
  ylab("R")+
  theme(legend.title = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x=element_blank(),
        legend.text = element_text(size=16),
        strip.text = element_text(size=16),
        axis.title.y = element_text(size=16),
        axis.text.y= element_text(size=14))+
  scale_alpha_manual(values = c(0.33,0.66, 1))

ggsave(paste0('pred_fold_cor_comparison.', out_device), device = out_device, path = fig_path, width = 20, height = 9, units = "in")
```

# $R^2$ collaped across folds

```{r}
iv_dfs = c("ez_t1_522_fa_3_condition_scores", "ez_t1_522_fa_3_condition_scores_t2subs", "ez_t1_522_fa_3_condition_scores_nont2subs")
dv_dfs = c('demog_fa_scores_t1')

out = data.frame()

for(iv_data in iv_dfs){
  for(dv_data in dv_dfs){
    file_name = paste0(input_path,'prediction/pred_out_', iv_data,'_', dv_data, '.csv')
    
    d = read.csv(file_name) 
    
    out = out %>%
      bind_rows(d)
  }
}
```

```{r}
out %>%
  mutate(iv = factor(iv, levels = c("drift_rate", "threshold", "non_decision"), labels=c("drift rate", "threshold", "non-decision")),
         dv = factor(dv, levels = c('Drug_Use','Mental_Health','Problem_Drinking','Daily_Smoking','Binge_Drinking','Obesity','Lifetime_Smoking','Unsafe_Drinking','Income_LifeMilestones'), labels = c('Drug Use','Mental Health','Problem Drinking','Daily Smoking','Binge Drinking','Obesity','Lifetime Smoking','Unsafe Drinking','Income/Life Milestones')),
         iv_data = factor(iv_data, levels = c("ez_t1_522_fa_3_condition_scores_t2subs", "ez_t1_522_fa_3_condition_scores_nont2subs", "ez_t1_522_fa_3_condition_scores"), labels = c("150","372","522"))) %>%
  ggplot(aes(iv, all_folds_r))+
  geom_bar(stat="identity", position=position_dodge(width=0.9), aes(alpha=iv_data, fill=iv))+
  facet_grid(~dv,labeller = label_wrap_gen(width = 2, multi_line = TRUE))+
  xlab("")+
  # ylab(expression(R^{2}))+  
  ylab("R")+  
  theme(legend.title = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x=element_blank(),
        legend.text = element_text(size=16),
        strip.text = element_text(size=16),
        axis.title.y = element_text(size=16),
        axis.text.y= element_text(size=14))+
  scale_alpha_manual(values = c(0.33,0.66, 1))
  
```
